---
title: "Permutation Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(coin)
library(pander)
```

### Background and Question

The data that is presented in the link below is collected from marriage records in Mobile County, Alabama.

##### <a href="javascript:showhide('dataset')">The DataSet </a>

<div id="dataset" style="display:none;">

```{r}
knitr::kable(Marriage)
```

</div>

The question that we are pursuing is as follows:
Is there a difference in ages between spouses?

###Paired Analysis

For this analysis there is a lot of coding. You can find the coding that has been performed by clicking the Code box that you find right below this arrow ------------------------------------------------>

```{r}
bride <- subset(Marriage, person == "Bride", select = c("bookpageID", "person", "age"))
groom <- subset(Marriage, person == "Groom", select = c("bookpageID", "person", "age"))
Marriage2 <- merge(bride, groom, by="bookpageID")

t.obs <- mean(Marriage2$age.x - Marriage2$age.y)

N <- 2000
t.sim <- rep(NA,N)
for (i in 1:N){
  t.sim[i] <- mean(  (Marriage2$age.x - Marriage2$age.y) * sample(c(1,-1), 49, replace = TRUE) )
}
#hist(t.sim)
#abline(v = t.obs)
pval <- (  sum(t.sim <= t.obs) + 1)/N * 2

```

```{r}
hist(Marriage2$age.x - Marriage2$age.y, xlab="GroomAge - BrideAge", main="Marriage Age Difference", col="green4")
```

The above histogram is a representation of the difference in ages between the bride and groom in Mobile County. 

Through our analysis and doing the permutation test we were able to calculate a p-value of 0.002.

###Interpretation

With our original question of having a difference in ages between spouses, we have been able to come up with a conclusion. The p-value of the test is 0.002 which is below our alpha level of 0.05. This means that we have sufficient evidence to reject the null, meaning that there is a difference in the ages between spouses.


