---
title: "Reading Comprehension"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(pander)
```

<br />

### Background

"This study investigated the effectiveness of explicit instruction in 'think aloud' as a means to promote elementary students' comprehension monitoring abilities" [(Baumann & Jones, 1992)](http://jlr.sagepub.com/content/24/2/143.full.pdf).

The following explanation of the experiment comes directly from the original research article [(Baumann & Jones, 1992)](http://jlr.sagepub.com/content/24/2/143.full.pdf).

##### <a href="javascript:showhide('uniquename')">The Experiment </a>

<div id="uniquename" style="display:none;">

<div style="padding-left:30px; padding-right:30px;">

Fourth-grade students from an elementary school in a rural mid-western community participated in the study. The school had adopted a team teaching plan, and three fourth-grade teachers shared an open classroom area. The teachers grouped the students for reading instruction across the three classes according to reading ability. One teacher each worked with a high [ability reading group], [one teacher worked with a] middle [ability group], and [the other teacher worked with the] low reading [ability] group. All groups used basal reading materials and were formed at the beginning of the school year on the basis of standardized achievement test data, prior performance in the basal reading program, and teacher judgment.

Sixty-eight of the 72 students in the fourth-grade team participated in the study; the four nonparticipating students were mainstreamed special education students for whom the instructional materials were considered to be too frustrating to read. The 68 students were assigned randomly to one of the three experimental groups (Strat = 23 students; DRTA = 23 students; Basal = 22 students). Because of illness and transfer to another school, incomplete data were obtained for one subject each from the Strat and DRTA groups. Thus, the final sample consisted of 66 students, 32 girls and 34 boys.

Posttests 1-3 were group-administered, quantitative measures obtained
for all students in the sample. Posttest 1 was an error detection test. Posttest 2 was a comprehension monitoring questionnaire. Posttest 3 was a modified cloze test. To statistically account for students'
preexperimental differences in comprehension monitoring abilities, two pretests were constructed, administered, and used as covariates in data analyses for Posttests 1-3.

Pretest 1. The first pretest involved an error detection task designed to evaluate students' preintervention ability to monitor their comprehension.... Students were provided the following directions which were printed on the pretest and read aloud by the experimenter: 'Read the story carefully. Try to understand it as you read. There will be 16 sentences that do not belong in the story. Look for them as you read. Ask yourself, "What does not make sense in this story?" Underline the 16 sentences that do not make sense. Check your work when you are finished. If you cannot read a
word, raise your hand, and we will help you.' Preceding Pretest 1, students were given a brief practice exercise that required them to identify obvious intruded sentences in two short texts. Pretest 1 was scored by calculating the number of correctly identified intruded sentences.

Pretest 2. The second pretest queried students about the strategies they believed to be useful in promoting their understanding of stories. In this pretest, modeled after an instrument developed by Schmitt (1988, 1990), the students were presented with 15 multiple-choice items like the following:

<div style="padding-left:15px;">
When I read, it is a good idea to:

A. sound out words I don't know.

B. make some guesses about what will happen in the story.

C. make a list of all the details in the story.

D. look up new words in the dictionary.

</div>

Posttest 1: Error detection test. Posttest 1 was the same in form as Pretest 1: with 16 intruded sentences lexically consistent at a local level but semantically inconsistent at a global level. 

Posttest 2: Comprehension monitoring questionnaire. Posttest 2, consisting of 18 items, was a slightly expanded version of Pretest 2. 

Posttest 3: Degrees of Reading Power. The Degrees of Reading Power (DRP)
(1986) is a standardized, commercially produced instrument that employs a modified cloze format (words are deleted from a selection and for each omission the student selects the most appropriate word from a list of five). According to the authors, DRP tests "are holistic measures of how well the messages within text are understood. . . . DRP tests focus measurement on determining how well students process or construct meaning from paragraphs as they read through a selection" (DRP Handbook, 1986, p. 1). Thus, the DRP, at one level, is a general reading comprehension assessment tool.

</div>

The data for this study are contained in the `Baumann` dataset.

###Question and Hypothesis

Given the data and having interest in teaching, I have decided to look at the performance of the students. Whenever we are looking at testing individual's performance, we wand to maximize the potential of those individuals. The question used for this analysis is, Which method provides the best results for comprehension? In order to address this question I chose to use the third post test as my measuring point. The third test is the Degrees of Reading Power (DRP) (1986) is a standardized, commercially produced instrument that employs a modified cloze format (words are deleted from a selection and for each omission the student selects the most appropriate word from a list of five). According to the authors, DRP tests “are holistic measures of how well the messages within text are understood. . . . DRP tests focus measurement on determining how well students process or construct meaning from paragraphs as they read through a selection” (DRP Handbook, 1986, p. 1). Thus, the DRP, at one level, is a general reading comprehension assessment tool.

The null hypothesis is that all groups are samples from the same distribution. The alternative hypothesis is that one of the groups will be a sample from a different distribution.

Alpha, or the level of significance, for this test is 0.05.

###Data

The data that follows shows the experiment and the tests that were administered.

##### <a href="javascript:showhide('dataset')">The DataSet </a>

<div id="data-set" style="display:none;">

```{r}
knitr::kable(Baumann)
```

</div>

```{r}
boxplot(post.test.3 ~ group, data=Baumann, type=c("p","a"), col='orange', xlab="Group", ylab="DRP", main="Post Test 3")

pander(favstats(post.test.3 ~ group, data=Baumann))

```


With this data I went forward and performed a Kruskal-Wallis Rank Sum Test based on the observations of the third post test, measuring the DRP. The groups given are Basal, which is the control, and then the two alternative styles DRTA and Strat.

```{r}

pander(kruskal.test(post.test.3 ~ group, data=Baumann))

```

The Kruskal-Wallis test that was run was able to show a lot of valuable information. The chi-squared test statistic of the Kruskal-Wallis test is 9.632. Given the fact that we have three different groups, we have two degrees of freedom. Lastly, the obtained p-value is 0.0081 which gives us sufficient evidence to make an assumption.

###Interpretation

Given the obtained p-value of 0.0081 we have sufficient evidence to reject the null hypothesis that at least one of the three groups produces different results. In other words, because our p-value is lower than 0.05, we can conclude that one of the groups comes from a different distribution. As we look at the following graph we are able to see that the differences may not be huge, but there is a difference. The Basal group has a lower median score than the other two teaching styles. DRTA has the highest performance, but at the same time, there are two students who's performance is lower than all other students in all three groups.

```{r}

xyplot(post.test.3 ~ group, data=Baumann, pch=16, col='orange', vertical=TRUE, xlab="Group", ylab="Degrees of Reading Power", main="Post Test 3 Results", type=c("p","a"))

```

I think that it would be interesting to do this experiment again but with larger numbers. Acknowledging the fact that we all have different of methods of learning, some exceptions might come with small groups like this.







